\bf Visual Diagnostics of Dynamic Social Network Models

### 2019-07-02 paper.Rnw
ditch senate data example, clarity in notation, some other minor edits, merge table 1 & figure 4

\documentclass[12pt]{article}
% \usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
% \geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\usepackage{graphicx}
\usepackage[font=small,skip=5pt]{caption}
\usepackage{subcaption}
\usepackage{animate}
\usepackage{afterpage}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
% \usepackage{color}
\usepackage{multirow}
\usepackage{rotating}
\usepackage[dvipsnames,svgnames,table]{xcolor}
\usepackage{hyperref}
\graphicspath{{figure/}}
%\usepackage{endfloat} % Figures to the end of the document

@@

\bigskip
\begin{abstract}
Statistical models for networks have been studied for decades, and they have gotten more complex as computing power has grown. Some network models, such as those for dynamic social networks networks, are analytically intractable and therefore require extensive simulation procedures for model fitting. In this paper, we use model visualization techniques introduced in Wickham et al (2015) to better understand some of these models and their behavior. We use a three-pronged visualization attack: (a) network visualizations to view the \emph{model}, (b) visualizations of collections of models to interpret parameters, and (c) visualizations of fitting algorithms to gain insights into the underlying mechanisms. With the help of static and dynamic visualizations, we bring the hidden algorithmic fitting processes into the foreground, with the goal of a better understanding and higher accessibility of statistical models for social network analysts.
\end{abstract}

>>>>

\documentclass[12pt]{article}
% \usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
% \geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\usepackage{graphicx}
\usepackage[font=small,skip=5pt]{caption}
\usepackage{subcaption}
\usepackage{animate}
\usepackage{afterpage}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
% \usepackage{color}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{rotating}
\usepackage[dvipsnames,svgnames,table]{xcolor}
\usepackage{hyperref}
\graphicspath{{figure/}}
\usepackage{endfloat} % Figures to the end of the document

@@

\bigskip
\begin{abstract}
Statistical models for networks have been studied for decades, and they have gotten more complex as computing power has grown. Some network models, such as those for dynamic social networks, are analytically intractable and therefore require extensive simulation procedures for model fitting. In this paper, we use model visualization techniques to better understand some of these models and their behavior. We use a three-pronged visualization attack: (a) network visualizations to view the model, (b) visualizations of collections of models to interpret parameters, and (c) visualizations of fitting algorithms to gain insights into the underlying mechanisms. With the help of static and dynamic visualizations, we bring the hidden algorithmic fitting processes into the foreground, with the goal of a better understanding and higher accessibility of statistical models for social network analysts.
\end{abstract}

### 2019-07-02 supporting-info-child.Rnw
ditch senate data example, clarity in notation, some other minor edits, merge table 1 & figure 4

%Outdegree and reciprocity have the same inverse relationship for both data sets. For the friendship data, the inclusion of $\\beta_3$ has strong effect on the estimates of the other two parameters, but not for the senate data.
<<compareModels, fig.cap="Density plots of objective function parameter estimates from repeatedly fitting models M1, M2, and M3 to the example data. ", fig.height = 3>>=
simu2$data <- "friends"
senateEsts$data <- "senate"
allDataEsts <- rbind(senateEsts, simu2)
allDataEsts$parameter <- as.factor(allDataEsts$parameter)
allDataEsts$parameter <- ordered(allDataEsts$parameter, levels = c(paste0("alpha", 1:3),paste0("beta", 1:4)))
ggplot(allDataEsts %>% filter(str_detect(parameter, "beta"))) +
  geom_density(aes(x = estimate, fill = Model), alpha = .5) +
  facet_grid(data~parameter, scales = "free", labeller = "label_both") +
  scale_fill_brewer(palette = "Greys") + 
  ThemeNoNet +
  theme(legend.position = 'bottom')
@

The density plots in Figure~\ref{fig:compareModels} show the distribution of the parameter estimates for the objective function parameters in M1, M2, and M3 the two example data sets we used. The most notable difference is in the values of $\beta_1, \beta_2$. In the senate data, the inclusion of $\beta_3$ has no effect on the estimates of $\beta_1, \beta_2$, while the opposite is true for the friendship data.

>>>>

%Outdegree and reciprocity have the same inverse relationship for both data sets. For the friendship data, the inclusion of $\\beta_3$ has strong effect on the estimates of the other two parameters, but not for the senate data.
<<compareModels, fig.cap="Density plots of objective function parameter estimates from repeatedly fitting models M1, M2, and M3 to the example data. ", fig.height = 3, eval=FALSE>>=
simu2$data <- "friends"
senateEsts$data <- "senate"
allDataEsts <- rbind(senateEsts, simu2)
allDataEsts$parameter <- as.factor(allDataEsts$parameter)
allDataEsts$parameter <- ordered(allDataEsts$parameter, levels = c(paste0("alpha", 1:3),paste0("beta", 1:4)))
ggplot(allDataEsts %>% filter(str_detect(parameter, "beta"))) +
  geom_density(aes(x = estimate, fill = Model), alpha = .5) +
  facet_grid(data~parameter, scales = "free", labeller = "label_both") +
  scale_fill_brewer(palette = "Greys") + 
  ThemeNoNet +
  theme(legend.position = 'bottom')
@

% The density plots in Figure~\ref{fig:compareModels} show the distribution of the parameter estimates for the objective function parameters in M1, M2, and M3 the two example data sets we used. The most notable difference is in the values of $\beta_1, \beta_2$. In the senate data, the inclusion of $\beta_3$ has no effect on the estimates of $\beta_1, \beta_2$, while the opposite is true for the friendship data.
